Check whether this version is the final version.
Which variables are we using?

```{r}
#Load libraries
#library(rgdal)
#library(maptools)
#library(raster)
library(rgeos)
library(tmap)
library(dplyr)
library(tidyr)
#install.packages("GISTools")
library(ggplot2)
library(sp)
library(car)
library(sf)
#install.packages("corrplot")
library(corrplot)
#install.packages("car")
library(ENVS450)
library(tidyverse)
library(stats)
```

```{r}
getwd()
```

```{r}
rm (list=ls( ))
```


```{r}
#Load shapefile and csv
LSOA_pop <- st_read(".", "LSOA_pop")
commute <- read.csv("LCR_methodTraveltoWork_LSOA.csv")
```


```{r}
#change data types
#LSOA$y2160 = as.numeric(as.vector(LSOA$X21.60y))
```

```{r}
#Check data types
#lapply(LSOA@data, class)
#str(LSOA$data)
```


```{r}
#check projection
st_crs(LSOA_pop)
```

### Commute and population variables
```{r}
#Calculate proportion of people that use bike
commute$perc_bike <- (commute$bicycle/commute$all) * 100
#Calculate proportion using public transport
commute$perc_ptrans <- ((commute$train_tube + commute$bus + commute$taxi_other)/commute$all)*100
#calculate proportion using car or motorbike
commute$perc_car <- ((commute$car_driver + commute$car_passenger + commute$motorbike)/commute$all)*100
```

```{r}
head(commute)
```


```{r}
#Subset commute csv
commute_bike <- commute[,c(1,2,14,15,16,17)]
#rename columns
colnames(commute_bike) <- c("LSOACD", "LSOA_name", "Slope", "Bike","PublicTransp", "Car" )
head(commute_bike)
```


```{r}
#calculate density
#LSOA_pop$density_night <- #LSOA_pop$all_ages/LSOA_pop$Area_ha
```

#### Population data
```{r}
#subset LSOA shapefile 
LSOA_pop <- LSOA_pop[,c(1:6,9,10,11,13,14,15)]
head(LSOA_pop)
lapply(LSOA_pop, class)
```

```{r}
LSOA_pop$Stn_bin <- as.numeric(paste(LSOA_pop$Stn_bin))
head(LSOA_pop)
```


```{r}
#join commute and LSOA shapefile
LSOA_bike_pop <- merge(LSOA_pop, commute_bike, by.x = "LSOA11CD", by.y = "LSOACD", all.x=TRUE)
head(LSOA_bike_pop)
```


#### IMD variable
```{r}
#read IMD csv
IMD <- read.csv("IMD_LCR_2015.csv")
head(IMD)
```

```{r}
#subset shapefile data
IMD <- IMD[,c(1,5)]
#rename columns
colnames(IMD) <- c("LSOACD", "IMD_Score")
```

```{r}
#Join IMD with LSOA
LSOA_IMD <- merge(LSOA_bike_pop, IMD, by.x ="LSOA11CD", by.y= "LSOACD", all.x=TRUE)
head(LSOA_IMD)
```

#### Green area and cycling variables
```{r}
#Load shapefiles
green <- st_read(".","LCR_GreenArea")
LSOA <- st_read(".", "LSOA_codes")
cycle_lane <- read_sf("cycle_final.shp")
```


```{r}
#change data types
#LSOA$y2160 = as.numeric(as.vector(LSOA$X21.60y))
```

```{r}
#Check data types
#lapply(LSOA@data, class)
#str(LSOA$data)
```


```{r}
#check projection
st_crs(green)
st_crs(LSOA)
st_crs(cycle_lane) # this is wgs84
```

```{r}
#Calculate green area per LSOA
#transform data from sp to sf
#green_sf <- st_as_sf(green)
#LSOA_sf <- st_as_sf(LSOA)

options(scipen = 999)
#intersect datasets and convert the result to tibble
intersection <- as_tibble(st_intersection(green, LSOA))
#intersection <- st_intersection(LSOA, green)

#add an area calculation of each green area intersected with LSOA  column to the tibble 
intersection$Area_Gr <- st_area(intersection$geometry)

#plot the layers
#plot (green$geometry, col='green')
#plot(LSOA$geometry, add=T)
#plot(intersection$geometry, col='red', add=T)

#group data by LSOA area and calculate the total greenspace area per LSOA
#output as new tibble
#area_LSOA <- intersection %>%
#  group_by(LSOA11CD) %>%
#  summarise(Area_Gr = sum(Area_Gr))
intersection <- aggregate(intersection$Area_Gr, by=list(LSOA11CD=intersection$LSOA11CD), FUN=sum)
#change column names
colnames(intersection) <- c("LSOA11CD", "Area_Gr")
#Change numbers to numeric
intersection$Area_Gr <- as.numeric(intersection$Area_Gr)



#join area_LSOA dataframe with LSOA shapefile
green_LSOA <- left_join(LSOA, intersection, by.x = "LSOA11CD", by.y = "LSOA11CD", all.x=TRUE)
```

```{r}
head(green_LSOA)
lapply(green_LSOA, class)
```


```{r}
#create proportion of green area per lsoa
green_LSOA$prop_green <- (green_LSOA$Area_Gr/green_LSOA$AreaM2)*100
```

```{r}
#Save shape
#writeOGR(Green_LSOAarea, ".", "LSOA_IMD", driver="ESRI Shapefile")
```

#### cycle lane variable

```{r}
#reproject
cycle_proj <- st_transform(cycle_lane, crs = st_crs(green))
```

```{r}
#Save reprojected cycling shape
#st_write(cycle_proj, "cycle_all_proj.shp", driver = "ESRI Shapefile")
```


```{r}
#check projection
st_crs(green)
st_crs(cycle_proj)
st_crs(LSOA)
```


#transform data from sp to sf
cycle_sf <- st_as_sf(green)
#LSOA_sf <- st_as_sf(LSOA)

https://gis.stackexchange.com/questions/280760/intersecting-lines-and-polygons-and-calculating-line-length-in-r

```{r}
#intersect cycle with LSOA and convert the result to tibble
intersection_cycle <- as_tibble(st_intersection(cycle_proj, LSOA))

#calculate length of cycle in each polygon
intersection_cycle$Length <- st_length(intersection_cycle$geometry)

#group data by LSOA and calculate the total cycle lane length per LSOA
length_LSOA <- intersection_cycle %>%
  group_by(LSOA11CD) %>%
  summarise(Length = sum(Length))
#change data type units of cycle lane length to numeric (from units with m suffix)
length_LSOA$Length <- as.numeric(length_LSOA$Length)
```

```{r}
#total length of cycling infrastructure in LCR
total_length <- sum(length_LSOA$Length)
```

Create a proportion of cycle lanes per LSOA with route cycleway added
Total of cycle lanes in all LCR is 736,832.3 mts/ 736.8km, mean 660.59m, range min max 0.082 - 21941.29m. Interquartile range 1st quartile 236.66, 3rd quartile 1377.66
Total cycle lane per Borough
Halton 175285.77m/ 175.28km
Knowlsey 62402.19/ 62.40 km
Liverpool 195203.43m/ 195.20km
Sefton 119563.78m / 119.56km
St Helens 62012.93m/ 62.01km
Wirral 122364.16m/ 122.36km

```{r}
#create proportion of cycle lane per lsoa
length_LSOA$prop_lane <- (length_LSOA$Length/total_length)*100
#merge with original LSOA data
green_cycle <-merge(green_LSOA, length_LSOA, by.x = "LSOA11CD", by.y = "LSOA11CD", all.x=TRUE)
head(green_cycle)
```



Old cycling lengths were calculated without merging "route - bicycle"" from OSM. Old values and model is csv called LSOA_model_old.csv

Total of cycle lanes in all LCR is 352,991 mts, mean 247m, range min max 0.83 - 5368. Interquartile range 1st wuartile 28.8686, 3rd quartile 267.975
Total cycle lane per Borough
Halton 98048.18215
Knowlsey 33650.11874
Liverpool 57777.48235
Sefton 58115.4427
St Helens 42762.64236
Wirral 61872.82516


Green_cycle has green area and proportion of green, also has total length of cycling lane
```{r}
#convert sf to dataframe
green_cycledf <- green_cycle %>% st_set_geometry(NULL)
class(green_cycledf)
```


##### Merge all data to create the final model
```{r}
#merge with original LSOA data
LSOA_model <-full_join(LSOA_IMD, green_cycledf, by = "LSOA11CD")
head(LSOA_model)
```

```{r}
summary(LSOA_model$Length)
```

```{r}
#https://datacarpentry.org/R-genomics/04-dplyr.html
#sum cycling length per borough in m
total_cycle <- LSOA_model %>% 
  group_by(layer) %>%
  summarise(Length = sum(Length, na.rm = TRUE))
head(total_cycle)
```

Halton	175285.77		
Knowsley	62402.19	
Liverpool	195203.43			
Sefton	119563.79			
StHelens	62012.93			
Wirral	122364.16


```{r}
#Transform sf greencycle to dataframe or we get an unordered dataframe if exported as csv
#green_df <- green_cycle %>% st_set_geometry(NULL)
#class(green_df)
```


```{r}
#Save shape
st_write(LSOA_model_shp, "LSOA_model.shp", driver="ESRI Shapefile")
#Save as csv
#write.csv(green_cycle, file="green_cycle.csv")
#greenarea_lane <- read.csv("green_cycle.csv")
```


https://gis.stackexchange.com/questions/224915/extracting-data-frame-from-simple-features-object-in-r

https://stackoverflow.com/questions/26049011/r-programming-how-to-filter-then-sum
```{r}
#Transform sf LSOA_model_shp to dataframe or we get an unordered dataframe if exported as csv
model_df <- LSOA_model %>% st_set_geometry(NULL)
class(model_df)
```


Here, we are going to export the final LSOA model df to calculate proportion of cycle and calculate the total length of cycling roads per LSOA. 
Test whether proportion in relation to total cycling infrastructure in LCR is more significant than cycling infrastructure proportion in relation to each borough total cycling lanes.

First export as csv/shape

```{r}
head(LSOA_model)
ncol(LSOA_model)
```


```{r}
#Save as csv
#write.csv(LSOA_df, file="LSOA_model_all.csv")
#Load model data
#LSOA_model <-read.csv("LSOA_model.csv")
#Replace NA values
model_df[,19:22][is.na(model_df[,19:22])] <- 0
head(model_df)
```

```{r}
#create the summary of all variables
#lapply(2:ncol(LSOA_model), function(j) summary(LSOA_model[c(1, j)], measurevar = 2, groupvars = 1))
#Obtain the summary of all variables
for(i in 2:ncol(LSOA_model)){
    summary<-summary(LSOA_model, measurevar = LSOA_model[,i], groupvars = LSOA_model[1])
}
head(summary)
```

```{r}
summary(LSOA_model)
```


##### Correlation and correlogram
Check correlation of variables



```{r}
#lapply(model_df, class)
#head(model_df)
```


```{r}
#subset lsoa to remove categorical variables
correlation_m <- cor(model_df[,c(5,6,7,13,14,15,16,17,20,22)])
#Rename columns
colnames(correlation_m) <- c("User","Station", "Density", "Slope","Bike","PublicTrans","Car","IMD","Green_Area","Lane")
#create correlation results
cor_results<- cor.results(correlation_m, sort.by = "abs.r", data = model_df, var.name = "Bike")
cor_results
head(correlation_m)
```



```{r}
#Load old LSOA model
model_old <- read.csv("LSOA_model_old.csv")
#subset lsoa to remove categorical variables
correlation_old <- LSOA_model[,c(5,9,10,12:16,19,21,22)]
#Rename columns
#colnames(correlation) <- c( "User","Density", "Slope","Bike","PublicTrans","Car","IMD","Green_Area","Lane_tot", "Lane_loc" )
#create correlation matrix
corr_old_m <- cor(correlation_old)
#create results df 
corr_old_res<- cor.results(corr_old_m, sort.by = "abs.r", data = correlation_old, var.name = "Bike")
corr_old_res
```



```{r}
#Check correlation of lane
cor_results_lane <- cor.results(correlation_m, sort.by = "abs.r", data = correlation_m, var.name = "Lane_tot")
cor_results_lane
```


```{r}
#create correlation matrix
corr_matrix <- cor(correlation)
#create correlogram
#corrplot(corr_matrix, type="upper", order = "hclust", col=brewer.pal(n = 8, name= "RdYlBu"), tl.col = "black", tl.srt = 45, tl.cex = .8, number.cex = .5)
```


```{r}
col <- colorRampPalette(c("#BB4444", "#EE9988", "#FFFFFF", "#77AADD", "#4477AA"))
plot <- corrplot(correlation_m, method = "color", col = col(200),
         type = "upper", order = "hclust", number.cex = .7,
         addCoef.col = "black", # Add coefficient of correlation
         tl.col = "black", tl.srt = 90, tl.cex = 0.7,
         # Text label color and rotation
         # Combine with significance
         #p.mat = p.mat, sig.level = 0.01, insig = "blank", 
         # hide correlation coefficient on the principal diagonal
         diag = FALSE)
```
The sign of the coefficient indicates the direction of the relationship. If both variables tend to increase or decrease together, the coefficient is positive, and the line that represents the correlation slopes upward. If one variable tends to increase as the other decreases, the coefficient is negative, and the line that represents the correlation slopes downward.


Correlation coefficients - definition
A correlation between variables indicates that as one variable changes in value, the other variable tends to change in a specific direction. A correlation coefficient measures both the direction and the strength of this tendency to vary together.

-A positive correlation indicates that as one variable increases the other variable tends to increase.
-A correlation near zero indicates that as one variable increases, there is no tendency in the other variable to either increase or decrease.
-A negative correlation indicates that as one variable increases the other variable tends to decrease.

Correlation
Car has a strong negative correlation with public transport, IMD and bike usage. This correlation represents that as car usage increase, ublic transport, bike usage and IMD tend to decrease. In contrast, when pulic transport usage increase, so does IMD score in an area. For bike usage, the strongest positive correlation is with density of population and the created cyclists profile. In this vein, cycle lane does not display strong correlation with cycling

```{r}
#corrplot(corr_m, method="number", type="lower", bg="pink")
```


```{r}
#corrplot.mixed(corr_m, lower.col = "black", number.cex = .7, tl.cex = .5, tl.col="orange" )
```



```{r}
#activate scientific notation
options(scipen = 0)
#options(digits = 5)
```

### Create the model for the entire LCR
#### Joint model
```{r}
#subset model_df to required columns
model_df <- model_df[,c(1,2,3,5,6,7,12:17,20,22)]
#rename column names in LSOA model
colnames(model_df) <- c("LSOACD", "all_ages","layer", "User_prop", "Stn_bin", "Density", "LSOA_name", "Slope", "Bike", "PublicTransp", "Car", "IMD_Score", "prop_green", "prop_lane")
```

```{r}
#Create the linear regression model, dependant variable perc bike
model1 <- lm('Bike  ~ Stn_bin + User_prop + Density + Slope + Car + PublicTransp + IMD_Score + prop_green + prop_lane', LSOA_model)
summary(model1)
vif(model1)
```
R-squared usually increases as variables increase. that is why AIC is reliable to check fo model

```{r}
skew(model_df$log(Bike))
```


```{r}
#check skew
ggplot(data=model_df) +
  geom_density( aes(x=Bike))
```

```{r}
model_df$logbike <- model_df$log1p(Bike)
```


```{r}
#FINAL MODEL
#http://rpubs.com/marvinlemos/log-transformation
#create another model for log of bikes and check coefficients
#Create the linear regression model, dependant variable perc bike
model1.log <- lm('log1p(Bike)  ~ User_prop + Stn_bin + Density + Slope + PublicTransp + IMD_Score + prop_green + prop_lane', model_df)
summary(model1.log)
```

```{r}
vif(model1.log)
```

```{r}
AIC(model1, model1.log)
```

```{r}
gvlma ::gvlma(model1.log)
```

```{r}
cbind(summary(model1)$coefficients,confint(model1))
```






#Standardize perc_car
https://stackoverflow.com/questions/15215457/standardize-data-columns-in-r

Standardize variables

zVar <- (myVar - mean(myVar)) / sd(myVar)

And makes using it dplyr much easier: mutate(var = (var - mean(var))/sd(var)). - RobertMyles Jun 2 '17 at 17:41
But can this be used to get the z-score for two variables? - lf_araujo Jul 11 '17 at 20:42
to denormalize myVar <- (zVar * sd(zVar)) + mean(zVar), right? - Artur_Indio Aug 14 '17 at 20:52 
2
@Artur_Indio Almost: newVar <- (zVar * sd(myVar)) + mean(myVar). You have to use the original mean/sd. As you wrote it, you'll multiply by sd(zVar)=1 and add mean(zVar)=0, so nothing will change :) - random_forest_fanatic Apr 24 '18 at 17:08
```{r}
#Standardize is substract the mean a and divide by standard deviation
LSOA_model$perc.car.st <- (LSOA_model$perc_car - mean(LSOA_model$perc_car)) / sd(LSOA_model$perc_car)

```

```{r}
head(LSOA_model)
```

```{r}
#Recreate model with transformed bike and car variables
#Rerun the model
model1.tr <- lm('perc_bike_tr  ~ Density + User_prop + Stn_bin+ perc_ptrans + perc.car.st+ IMD_Score + prop_lane + avslope_perc_u10km', LSOA_model)
summary(model1.tr)
```

```{r}
#Recreate model2 with transformed bike and remove car variable
model2.tr <- lm('perc_bike_tr  ~ Density + User_prop + Stn_bin+ perc_ptrans + IMD_Score + prop_lane + avslope_perc_u10km', LSOA_model)
summary(model2.tr)
```

```{r}
summary(model2)
```

```{r}
#remove the variable with the highest p-value, cycle-lane
model4 <- lm('perc_bike_tr  ~ Density + User_prop + Stn_bin+ perc_ptrans + IMD_Score  + avslope_perc_u10km', LSOA_model)
summary(model4)
```


```{r}
vif(model1)
```
model 1 woth all variables untransformed
model1_tr bike is transformed to achieve normality
model 2 car removed
model2-tr bike and car transformed
model 3 public transport removed
model 4 remove cycle lane
```{r}
AIC(model1, model1_tr, model2, model2.tr)
```

#Model is linear
```{r}
crPlots(model2)
```


#Checking for collinearity
create a model without any spatial data, export lsoamodel to csv and run them.
model 1 all variables
model 2 remove per car
model 3 remove per p transport
```{r}
vif(model1_tr)
```

```{r}
vif(model2)
```

```{r}
summary(model2)
```


```{r}
mean(vif(model1))
```



##Create a model for Liverpool
```{r}
#Subset LSOA_model for Liverpool only
model_liv <- model_df[model_df$layer == 'Liverpool',]
#Create model for Liverpool
Liv_model <- lm('Bike  ~ User_prop + Stn_bin+ Density + Slope + Car  + PublicTransp + IMD_Score + prop_green+ prop_lane', model_liv)
summary(Liv_model)
```

```{r}
#FINAL MODEL
#Create model for Liverpool log transformed prop bike
liv_log <- lm('log1p(Bike) ~ User_prop + Stn_bin + Density + Slope + Car + PublicTransp + IMD_Score + prop_green + prop_lane', model_liv)
summary(liv_log)
```


```{r}
gvlma ::gvlma(liv_log)
```


```{r}
#Subset LSOA_model for Wirral
model_wirr <- LSOA_model[model_df$layer == 'Wirral',]
```

# Model for Wirral
```{r}
#Create model for Wirral without station binary
Wirr_model <- lm('Bike  ~ User_prop + Density + Slope + PublicTransp + IMD_Score + prop_green + prop_lane', model_wirr)
summary(Wirr_model)
```

```{r}
vif(Wirr_model)
```

```{r}
#FINAL MODEL
#model bike log transformed
wirr_log <- lm('log1p(Bike)  ~ User_prop + Density + Slope+ PublicTransp+ IMD_Score + prop_green + prop_lane', model_wirr)
summary(wirr_log)
```

#### Checking for collinearity
.
```{r}
vif(wirr_log)
```

```{r}
rownames(LSOA_model)
```


#### Create the model for the other boroughs
```{r}
#Subset LSOA_model for Halton, Knowsley, Sefton, St Helens
model_boroughs <- subset(LSOA_model,layer=='Halton'|layer=='Knowsley'|layer=='Sefton'|layer=='StHelens')
#model_boroughs <- LSOA_model[LSOA_model$layer == 'Halton',]+ [LSOA_model$layer == 'Knowsley',]+ [LSOA_model$layer == 'Sefton',]+[LSOA_model$layer == 'St Helens',]
head(model_boroughs)
```

```{r}
#Create the model
#Create model for Wirral without station binary
borough_model <- lm('Bike  ~ User_prop + User + Density_day + density_night + Slope + PublicTransp + Car+ IMD_Score + Lane + prop_green', model_boroughs)
summary(borough_model)
```

```{r}
vif(borough_model)
```

