Envorinmental variables

```{r}
getwd()
```


```{r}
#install.packages("ncdf4")
library(raster)
library(ncdf4)
#library(dplyr)
```


```{r}
boundary <- shapefile("LCR_LSOAS.shp")
rain <- nc_open("rainfall_hadukgrid_uk_1km_ann_201701-201712.nc")
#st_crs(rain)
#st_crs(boundary)
```

```{r}
#create a text file with t he metadata
{
  sink('rainfall_hadukgrid_uk_1km_ann_201701-201712_metadata.txt')
  print(rain)
  sink()
}
```

```{r}
#get the list of the raster variable names
attributes(rain$var)$names
```

```{r}
#get units of variables
ncatt_get(rain, attributes(rain$var)$names[1])
```

https://stackoverflow.com/questions/46207167/how-to-extract-from-a-nc-file-based-on-a-shapefile-in-r


```{r}
rain_matrix <- ncvar_get(rain, varid="rainfall") # extact value. In this case, 3 is for time step number 3
rain_r <- raster(t(rain_matrix),xmn=-180, xmx=180, ymn=-90, ymx=90) # create a raster with data (it is flipped)
rain_r <- flip(rain_r,2) # correct raster
projection(rain_r) <- '+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0 ' # add CRS

plot(rain_r) # check result
```

```{r}
# reproject shapefile
g <- st_transform(boundary, crs = st_crs(rain_r@crs))
#g <- spTransform(boundary, rain_r@crs)
# extract values by each feature
rain_values <- extract(rain_r,g, fun=mean,na.rm=T, df=TRUE)
#extrac and add as new column in shapefile
#boundary$rain <- extract(rain_r, g, na.rm = TRUE)
# extract mean by each feature
#rain_mean <- extract(rain_r,g,fun=mean,na.rm=T, df=TRUE)
rain_values <- as.data.frame(rain_values)
#rain_mean$id <- seq.int(nrow(rain_mean))
#rain_mean
```

rain_mean in the mean annual rain per LSOA


```{r}
#create an id column for LSOAs and join
boundary$ID <- seq.int(nrow(boundary))
rain_LSOA <-full_join(boundary, rain_mean, by="ID")
#st_write(rain_LSOA, "rain_LSOA.shp", driver = "ESRI Shapefile")
colnames(rain_LSOA)
colnames(rain_LSOA) <- c("LSOA11CD", "spjoin_rif","ID_rain","Rain","id","geometry" )
```


#extract values for mean air temp
```{r}
#load raster
mean <- nc_open("tas_hadukgrid_uk_1km_ann_201701-201712.nc")

#get the list of the raster variable names
attributes(mean$var)$names
```

```{r}
#get units of variables
ncatt_get(mean, attributes(max$var)$names[1])
```

```{r}
# extract values
# extract values
mean_matrix <- ncvar_get(mean, varid="tas")
# create a raster with data (it is flipped)
mean_air <- raster(t(mean_matrix),xmn=-180, xmx=180, ymn=-90, ymx=90) 
mean_air <- flip(mean_air,2) # correct raster
# add CRS
projection(mean_air) <- '+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0 ' 
# check result
plot(mean_air) 
```

```{r}
# reproject shapefile
g <- spTransform(boundary, mean_air@crs)
#g <- st_transform(boundary, crs = st_crs(mean_air@crs))
# extract values by each feature
meanair_values <- extract(mean_air,g, fun=base::mean, na.rm=T, df=TRUE)
#extrac and add as new column in shapefile
#boundary$rain <- extract(rain_r, g, na.rm = TRUE)
# extract mean by each feature
#rain_mean <- extract(rain_r,g,fun=mean,na.rm=T, df=TRUE)
meanair_values <- as.data.frame(meanair_values)
#maxair_values$ID <- seq.int(nrow(minair_values))
```


```{r}
library(dplyr)
#create an id column for LSOAs and join
boundary$ID <- seq.int(nrow(boundary))
mean_LSOA <-merge(boundary, meanair_values, by="ID")
#st_write(rain_LSOA, "rain_LSOA.shp", driver = "ESRI Shapefile")
colnames(max_LSOA)
colnames(max_LSOA) <- c("LSOA11CD", "spjoin_rif","ID","max","geometry" )
```

```{r}
library(rgdal)
library(sf)
writeOGR(mean_LSOA, ".", "meanair_LSOA.shp", driver = "ESRI Shapefile")
#st_write(mean_LSOA, "meanair_LSOA.shp", driver = "ESRI Shapefile")

```


#### Do the same for min air temp
```{r}
min <- nc_open("tasmin_hadukgrid_uk_1km_ann_201701-201712.nc")
```

```{r}
#create a text file with t he metadata
{
  sink('tasmin_hadukgrid_uk_1km_ann_201701-201712.txt')
  print(min)
  sink()
}
```

```{r}
#get the list of the raster variable names
attributes(min$var)$names
```

```{r}
#get units of variables
ncatt_get(min, attributes(min$var)$names[1])
```

```{r}
# extract values
min_matrix <- ncvar_get(min, varid="tasmin")
# create a raster with data (it is flipped)
min_air <- raster(t(min_matrix),xmn=-180, xmx=180, ymn=-90, ymx=90) 
min_air <- flip(min_air,2) # correct raster
# add CRS
projection(min_air) <- '+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0 ' 
# check result
plot(min_air) 
```

```{r}
# reproject shapefile
g <- st_transform(boundary, crs = st_crs(min_air@crs))
# extract values by each feature
minair_values <- extract(min_air,g, fun=mean, na.rm=T, df=TRUE)
#extrac and add as new column in shapefile
#boundary$rain <- extract(rain_r, g, na.rm = TRUE)
# extract mean by each feature
#rain_mean <- extract(rain_r,g,fun=mean,na.rm=T, df=TRUE)
minair_values <- as.data.frame(minair_values)
#minair_values$ID <- seq.int(nrow(minair_values))
```

```{r}
#create an id column for LSOAs and join
boundary$ID <- seq.int(nrow(boundary))
min_LSOA <-full_join(boundary, minair_values, by="id")
#st_write(rain_LSOA, "rain_LSOA.shp", driver = "ESRI Shapefile")
colnames(min_LSOA)
colnames(min_LSOA) <- c("LSOA11CD", "spjoin_rif","ID","min","geometry" )
```


#extract values for max air
```{r}
#load raster
max <- nc_open("tasmax_hadukgrid_uk_1km_ann_201701-201712.nc")

#get the list of the raster variable names
attributes(max$var)$names
```

```{r}
#get units of variables
ncatt_get(max, attributes(max$var)$names[1])
```

```{r}
# extract values
# extract values
max_matrix <- ncvar_get(max, varid="tasmax")
# create a raster with data (it is flipped)
max_air <- raster(t(max_matrix),xmn=-180, xmx=180, ymn=-90, ymx=90) 
max_air <- flip(max_air,2) # correct raster
# add CRS
projection(max_air) <- '+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0 ' 
# check result
plot(max_air) 
```

```{r}
# reproject shapefile
g <- st_transform(boundary, crs = st_crs(min_air@crs))
# extract values by each feature
maxair_values <- extract(max_air,g, fun=mean, na.rm=T, df=TRUE)
#extrac and add as new column in shapefile
#boundary$rain <- extract(rain_r, g, na.rm = TRUE)
# extract mean by each feature
#rain_mean <- extract(rain_r,g,fun=mean,na.rm=T, df=TRUE)
maxair_values <- as.data.frame(maxair_values)
#maxair_values$ID <- seq.int(nrow(minair_values))
```


```{r}
#create an id column for LSOAs and join
boundary$ID <- seq.int(nrow(boundary))
max_LSOA <-full_join(boundary, maxair_values, by="ID")
#st_write(rain_LSOA, "rain_LSOA.shp", driver = "ESRI Shapefile")
colnames(max_LSOA)
colnames(max_LSOA) <- c("LSOA11CD", "spjoin_rif","ID","max","geometry" )
```

We have three LSOA shapefiles with rain, min and max. We need to join into one using LSOA codes 
```{r}
#join commute and LSOA shapefile
#rain.min<- st_join(rain_LSOA, min_LSOA)
#head(rain.min)
#join rain and min
rain.min <-full_join(rain_LSOA, min_LSOA, by="ID")
```

```{r}
st_write(rain_LSOA, "rain_LSOA.shp", driver= "ESRI Shapefile")
st_write(min_LSOA, "min_LSOA.shp", driver= "ESRI Shapefile")
st_write(max_LSOA, "max_LSOA.shp", driver= "ESRI Shapefile")
```




https://stackoverflow.com/questions/22333473/how-do-i-extract-raster-values-from-polygon-data-then-join-into-spatial-data-fra

e <- extract(r, p)
ee <- data.frame(ID=p$Block_ID, e)
```{r}
# reproject shapefile from projected to WGS84
#g <- st_transform(boundary, st_crs(rain_r@crs))
#extract and add as new column in shapefile
boundary$maxair <- extract(min_air, g, fun = mean, na.rm = TRUE)
```

```{r}
library(rgdal)
writeOGR(boundary, "env_variable.shp", driver= "ESRI Shapefile")
```


```{r}
# reproject shapefile from projected to WGS84
#g <- st_transform(boundary, st_crs(rain_r@crs))
#extract and add as new column in shapefile
boundary$minair <- extract(min_air, g, na.rm = TRUE)
```


# read the netcdf file as raster
nc_raster = raster("input_file.nc")

# if raster crs is NA, set the correct projection using proj4
projection(nc_raster) = CRS(+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0)

# save output
writeRaster(nc_raster, "nc_raster.tif", "GTiff")



```{r}
#test load and save as tiff
# read the netcdf file as raster
nc_maxair = raster("tasmax_hadukgrid_uk_1km_ann_201701-201712.nc")

# if raster crs is NA, set the correct projection using proj4
projection(nc_maxair) <- '+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0'

# save output
writeRaster(nc_maxair, "tasmaxair.tif", "GTiff")

#Load raster
maxair <- raster("tasmaxair.tif")
```
























```{r}
#this command is for normal rasters tiff not ncdf4
test_max <- extract(max_air, boundary, df=TRUE) 
```


centroids = st_transform(centroids, st_crs(liv.shape))

```{r}
#reproject to osgb36
maxtemp_pr = st_transform(maxtemp, st_crs(boundary))
st_crs(maxtemp_pr)
```

```{r}
st_write(maxtemp_pr, "maxtemp.shp", driver = "ESRI Shapefile")
```

```{r}
plot(maxtemp_pr$geometry, axes=TRUE)
plot(boundary$geometry, add=TRUE)
```

