
```{r}
#install.packages("stplanr")
library(stplanr)
library(dplyr)
library(rgdal)
library(sf)
library(tidyverse)
library(pct)
library(tmap)
#Saving  my API token to R environment
mytoken = 'd6fee7f1acac5c7c'
Sys.setenv(CYCLESTREET = mytoken)
```

```{r}
getwd()
```



```{r}
#Add LCR zone polygon with pct tool
liv_shape <- pct_regions[pct_regions$region_name=="liverpool-city-region", ]
plot(liv_shape)
```


```{r}
#add LCR boundary
#Add MSOA LCR centroids
boundary <- st_read("LCR_Boundary.shp")
boundary = st_transform(boundary, st_crs(liv_shape))
#LCR_cent$lsoa11cd <- as.character(LCR_cent$lsoa11cd)
plot(boundary$geometry)
#head(boundary)
```

```{r}
#Extract some data
od = stplanr::od_data_sample
class(od)
od
```

```{r}
getwd()
```

```{r}
rm (list=ls( ))
```

```{r}
#Load Od table
OD_LCR <- read.csv("LCR_LSOA_OD.csv")
OD_LCR <- OD_LCR[,c(2,4,5)]
names(OD_LCR) <- c("origin", "destination", "trips")
head(OD_LCR)
```

```{r}
#Convert to tibble
OD_LCR.t <- as_tibble(OD_LCR)
head(OD_LCR.t)
#convert to character
#OD_LCR$LSOACD <- as.character(OD_LCR$LSOACD)
#OD_LCR$LSOACD.1 <- as.character(OD_LCR$LSOACD.1)
#head(OD_LCR)
```


```{r}
#inter zones, regional total flows
(inter = OD_LCR %>% filter(LSOACD != LSOACD.1))
(intra = OD_LCR %>% filter(LSOACD == LSOACD.1))
```

### This section is about creating the black and white map for the cover of the dissertation.
All these chunks of code are based on this post from James Cheshire and his amazing map on the book "London: the information capital". https://jcheshire.com/visualisation/mapping-flows/


```{r}
#Add MSOA LCR centroids
LCR_cent <- st_read("LCR_weigthted_centroids.shp")
LCR_cent = st_transform(LCR_cent, st_crs(liv_shape))
head(LCR_cent)
LCR_cent$lsoa11cd <- as.character(LCR_cent$lsoa11cd)
plot(LCR_cent$geometry)
```


```{r}
#add LSOA centroids all
#cent_LSOA <- st_read("Lower_Layer_Super_Output_Areas_December_2011_Population_Weighted_Centroids.shp")
#Project
cent_LSOA = st_transform(LCR_cent, st_crs(liv_shape))
#Transform code from factor to character
#cent_LSOA$lsoa11cd <- as.character(cent_LSOA$lsoa11cd)
#Plot geometry
plot(liv_shape)
plot(cent_LSOA$geometry)
```

```{r}
#Clip centroids to Liv shape
cents_LSOA.liv <- cent_LSOA[liv_shape, ]
#Transform code from factor to character
cents_LSOA.liv$lsoa11cd <- as.character(cents_l$lsoa11cd)
plot(cents_LSOA.liv$geometry)
```

```{r}
#write.csv(cents_LSOA.liv, file="centroids_LSOA_LCR.csv")
cents.lsoa <- read.csv("centroids_LSOA_LCR.csv")
cents.lsoa$lsoa11cd <- as.character(cents.lsoa$lsoa11cd)
lapply(cents.lsoa, class)
```

```{r}
#Join centroid to od table
#merge origin xy coords
or_xy <- merge(OD_LCR, cents.lsoa, by.x="origin", by.y="lsoa11cd")
#change column names
names(or_xy) <- c("Origin", "Destination", "trips", "oX", "oY")
#merge destination xy cords
dest.xy <- merge(or_xy, cents.lsoa, by.x="Destination", by.y="lsoa11cd")
#change names
names(dest.xy) <- c("Origin", "Destination", "trips", "oX", "oY", "dX", "dY" )
```


```{r}
head(dest.xy)
```

```{r}
#Remove axes
xquiet<- scale_x_continuous("", breaks=NULL)
yquiet<-scale_y_continuous("", breaks=NULL)
quiet<-list(xquiet, yquiet)
```


```{r}
#Filter by excluding flows <10
map <- ggplot(dest.xy[which(dest.xy$trips>10),], aes(oX, oY))+
#The next line tells ggplot that we wish to plot line segments. The "alpha=" is line transparency and used below 
geom_segment(aes(x=oX, y=oY,xend=dX, yend=dY, alpha=trips), col="white",breaks= c(10,50,100,150))+
#Here is the magic bit that sets line transparency - essential to make the plot readable
scale_alpha_continuous(range = c(0.03, 0.3))+
theme(panel.background = element_rect(fill = 'black', colour='black'))+quiet+coord_equal()
map
```

```{r}
ggsave("map_LCR.png", plot = map, width = 15, height = 15, dpi = "screen")
```


```{r}
#This is the method to travel to work by LSOA for the LCR
OD_LCR.method <- get_pct_zones(region = "liverpool-city-region", geography = "msoa", extension = ".Rds")
```

```{r}
head(OD_LCR.method)
```


```{r}
#Step 1
# Load OD table for all UK
OD_UK <- pct::get_od()
#load MSOA centroids
centroids <- get_centroids_ew()
#set the same coordinates system
centroids = st_transform(centroids, st_crs(liv_shape))
#clip centroids for LCR
centroids.lcr <- centroids[liv_shape, ]
#filter flows within the lcr from the od uk
#od.LCR <- OD_UK %>%
#  filter(geo_code1 %in% centroids.lcr$msoa11cd)%>%
#  filter(geo_code2 %in% centroids.lcr$msoa11cd)
#subset flows with origin and destination within the LCR
od.LCR = OD_UK[
  OD_UK$geo_code1 %in% centroids.lcr$msoa11cd &
  OD_UK$geo_code2 %in% centroids.lcr$msoa11cd , 
]
```

Od_uk is the table with origin and destination per method of travel to work for the whole UK

```{r}
write.csv(centroids.lcr, file = "centroids_MSOA.csv" )
head(centroids.lcr)
```

```{r}
#convert from long to matrix
#MSOA_matrix <- od_to_odmatrix(lines.inter[1:3])
#write.csv(MSOA_matrix, file = "MSOA_matrix.csv")
```

In total, we have 26,774 trajectories. We first select inter regional 
```{r}
summary(od.LCR)
```

If we dont create the lines from od data, we could use the pct tools to extract the lines already. We need to compare the total number of lines when using pct_get_lines. This function gets the lines from the pct with all attributes

```{r}
lcr_odlines <- get_pct_lines(region = "liverpool-city-region", geography = "msoa")
nrow(lcr_odlines)
plot(lcr_odlines$geometry)
```

Continue using stplanr, create lines inter

```{r}
#Step 2
#create desire lines for LCR
lines.lcr <- od2line(od.LCR, centroids.lcr)
nrow(lines.lcr)
#create inter regional flows
lines.inter <- lines.lcr %>% filter(geo_code1 != geo_code2)
#create intra regional
lines.intra <- lines.lcr %>% filter(geo_code1 == geo_code2)
```
Total 26774
Inter region 26574
Intra MSOA 200

```{r}
#Save file
write.csv(lines.inter, file="lines_inter.csv")
```


```{r}
#Step 3
#To avoid plotting 26000 lines we need to subset
#select top all flows the mean is 17 and min is 1 max is 965
#select a new threshold for all 
trips_all_min <- 50
#Select all flows according to the minimum set
lines.top <- lines.inter %>% filter (all >= trips_all_min)
#select both inter and intra regional with the minimum set
#top.intra.inter <- lines.lcr %>% filter(all>= 30)
nrow(lines.top)
nrow(lines.lcr)



#create top all commutes
# 1 commute = total 2205 trips
#2 commutes =  782
#Set a threshold bof cycling
#bike.min <- 3
#bikes.top <- lines.inter %>% filter (bicycle >= bike.min)
```

By creating the trips min to 50, check how representative it is.
```{r}
#calculate the proportion covered by the top lines
nrow(lines.top)/ nrow(lines.lcr)
#calculate sum of total commmutes by each dataset
sum(lines.top$all) / sum(lines.lcr$all)
```
7% of the top inter intra flows cover 50% of all flows

```{r}
#calculate the proportion covered by the top lines
nrow(lines.top)/ nrow(lines.lcr)
#calculate sum of total commmutes by each dataset
sum(lines.top$all) / sum(lines.lcr$all)
```
12% of the top lines conver 60% of the total flows within the region.

```{r}
nrow(lines.top)
nrow(lines.lcr)
```


```{r}
summary(lines.top$bicycle)
```


```{r}
#plot inter regional flows
lwd = lines.top$all / mean(lines.top$all)/10
#create a percentage active
#lines.lcr$percent.active <- (lines.lcr$bicycle+lines.lcr$foot) / lines.lcr$all *100
#create percent that does not drive
lines.top$no_drive <- 100 - (lines.top$car_driver) / lines.top$all *100
#create percent not drive in the overall lines file
lines.lcr$no_drive <- (lines.lcr$car_driver) / lines.lcr$all *100
#create a percentage cycle
#lines.lcr$percent.cycle <- (lines.lcr$bicycle/ lines.lcr$all) *100
#plot the map
plot(lines.top["no_drive"], lwd = lwd)
```

```{r}
#summary(lines.top$no_drive)
head(lines.top)
```

```{r}
#plot all inter regional flows. Select most representative
lines.top = lines.top %>% 
  arrange(all) 
tm_shape(liv_shape) + tm_borders() +
  tm_shape(lines.top) +
  tm_lines(palette = "viridis", style = "quantile", n = 5,
    lwd = "all",
    scale = 9,
    title.lwd = "Number of trips",
    alpha = 0.5,
    col = "all",
    title = "Total flows",
    legend.lwd.show = FALSE) +
  tm_scale_bar(width = 0.08,lwd= 0.3, 
               position = c("left", "bottom")) +
  tm_layout(
    legend.title.size = 1,
    legend.bg.alpha = 0.5,
    legend.bg.color = "white",
    #legend.position = c("right", "top"),
    #legend.height = 0.5,
    legend.text.size = 0.6,
    legend.stack="vertical",
    #bg.color = "grey85",
    frame = TRUE,
    frame.lwd = 0.5,
    legend.show = TRUE,
    legend.width = 1
    )
```

```{r}
summary(top.intra.inter$all)
```

```{r}
nrow(top.intra.inter)
```


```{r}
#plot no drive both intra and inter regional flows
no_drive.inter <- tm_shape(liv_shape) + tm_borders() +
  tm_shape(top.intra.inter) +
  tm_lines(palette = "-viridis", style = "pretty", n = 5,
    lwd = "all",
    scale = 7,
    title.lwd = "Number of trips",
    alpha = 0.5,
    col = "no_drive",
    title = "No car flows (%)",
    legend.lwd.show = TRUE) +
  tm_scale_bar(size = 0.3, width = 0.15,lwd= 0.3,
               position = c("left", "bottom")) +
  tm_compass(type = "arrow", size = 0.7, position = c("right", "bottom"))+
  tm_layout(
    legend.title.size = 0.6,
    legend.bg.alpha = 0.8,
    legend.bg.color = "white",
    legend.position = c("right", "top"),
    legend.height = 0.5,
    legend.text.size = 0.5,
    legend.stack="vertical",
    frame = TRUE,
    frame.lwd = 0.5,
    legend.width = 1)
no_drive.inter
```

```{r}
lines.top$id <- paste(lines.top$geo_code1, lines.top$geo_code2, sep= " ")
```

#this line creates the routes for the top 5 rows of the data
street_lines <- line2route(lines_lcr[1:5, ], plan="quietest")

```{r}
#create cycling network
route_net <- SpatialLinesNetwork(rnet)
nodes_df <- data.frame(
    start = rep(c(1, 2, 3, 4, 5), each = 4),
    end = rep(c(50, 51, 52, 33), times = 100)
  )
rnet_bike_links <- sum_network_links(route_net, nodes_df)
plot(rnet_bike_links$geometry, col="red", lwd = 4, axes=TRUE)
plot(liv_shape$geometry, add=TRUE)
```



```{r}
#create cycle network
route_net <- SpatialLinesNetwork(rnet)
rnet_bike <- sum_network_routes(route_net, start = 1, end = 1500, sumvars = "length")
plot(rnet_bike, col="red", lwd = 4)
plot(route_net)
```

```{r}
head(route_net)
```


```{r}
#create street level lines from lines.top
route_lines <- line2route(lines.top[1:10, ], route_fun = route_osrm)
plot(liv_shape$geometry, axes=TRUE)
plot(route_lines$geometry, add= TRUE)
```


```{r}
tmap_save(no_drive.inter, filename = "no_drive_both.jpeg")
```


```{r}
nrow(lines.top)
```

  tm_borders(col = "grey", lwd = 0.3) + 
  tm_scale_bar(size = 0.3, width = 0.15 ,position = c("left", "bottom"))+
  tm_compass(type = "arrow", size = 0.7, position = c("right", "bottom"))+
  tm_layout(legend.position = c("right", "top"))
```{r}
#plot flows not by car
#lines.top = lines.top %>% 
 # arrange(no_drive) 
no_drive <- tm_shape(liv_shape) + tm_borders() +
  tm_shape(lines.top) +
  tm_lines(palette = "viridis", style = "pretty", n = 5,
    lwd = "all",
    scale = 9,
    title.lwd = "Number of trips",
    alpha = 0.5,
    col = "no_drive",
    title = "No car flows (%)",
    legend.lwd.show = TRUE) +
  tm_scale_bar(size = 0.3, width = 0.15,lwd= 0.3,
               position = c("left", "bottom")) +
  tm_compass(type = "arrow", size = 0.7, position = c("right", "bottom"))+
  tm_layout(
    legend.title.size = 0.6,
    legend.bg.alpha = 0.8,
    legend.bg.color = "white",
    #legend.position = c("right", "top"),
    #legend.height = 0.5,
    legend.text.size = 0.5,
    legend.stack="vertical",
    frame = TRUE,
    frame.lwd = 0.5,
    legend.show = TRUE,
    legend.width = 1)
no_drive
```

```{r}
  tmap_save(no_drive, filename = "noDrive_map.jpg")
```



```{r}
#Create same map but for car passenger and driver
#plot inter regional flows
lwd = lines.top$all / mean(lines.top$all)/10
#create a percentage car drivers
lines.top$driving<- 100 - (lines.top$car_driver+lines.top$car_passenger) / lines.top$all *100
#plot the map
plot(lines.top["driving"], lwd = lwd)
```

Exmaples 
```{r}
#tmap for commutes by car
lines.top = lines.top %>% 
  arrange(driving) 
tm_shape(liv_shape) + tm_borders() +
  tm_shape(lines.top) +
  tm_lines(palette = "plasma", style = "equal",
    lwd = "all",
    scale = 9,
    n=5,
    title.lwd = "Number of trips",
    alpha = 0.5,
    col = "driving",
    title = "Car flows (%)",
    legend.lwd.show = TRUE,
    legend.lwd.is.portrait = TRUE
  ) +
  tm_scale_bar() +
  tm_layout(
    legend.title.size = 1.5,
    legend.bg.alpha = 0.5,
    legend.bg.color = "white",
    legend.text.size = 0.8,
    legend.width=1.5,
    legend.stack="horizontal",
    legend.position = c("right", "top"))
  #tm_legend(outside=FALSE, stack="horizontal")
```


```{r}
#plot inter regional flows
lwd = lines.top$all / mean(lines.top$all)/10
#create a percentage active
#lines.lcr$percent.active <- (lines.lcr$bicycle+lines.lcr$foot) / lines.lcr$all *100
#create percent that does not drive
lines.top$no_drive <- 100 - (lines.top$car_driver) / lines.top$all *100
#create a percentage cycle
#lines.lcr$percent.cycle <- (lines.lcr$bicycle/ lines.lcr$all) *100
#plot the map
plot(lines.top["no_drive"], lwd = lwd, breaks= c(0,50,70,100,150,200))
```

```{r}
lines.inter$dist_bike <- st_length(lines.inter$bicycle)
```


## With this same dataset (lines.inter), create the map for top cycling flows

```{r}
summary(lines.inter$bicycle)
```

summaries
All flows mean 15.91, max 722
Bicycle 0.37, max 47
Car mean 9.257 max 451
Total LCR flows all 26775


```{r}
#Step 4= for all flows
#To avoid plotting 26000 lines we need to subset by bicycle
#select top all flows the mean is .37 and min is 0 max is 47
#select a new threshold for bicycle
trips_all_min <- 50
#Select all flows according to the minimum set
lines.top <- lines.inter %>% filter (all >= trips_all_min)
#select both inter and intra regional with the minimum set
#top.intra.inter <- lines.lcr %>% filter(all>= 30)
nrow(lines.top)
nrow(lines.lcr)
```

```{r}
nrow(lines.inter)
```


```{r}
#Step 4 = for bike flows
#To avoid plotting 26000 lines we need to subset by bicycle
#select top all flows the mean is .37 and min is 0 max is 47
#select a new threshold for bicycle
bike_min <- 3
#define minimum trips for inter flows
bike.top <- lines.inter %>% filter (bicycle >= bike_min)
nrow(bike.top)
#define minimum trips for all flows
#bike.top.all  <- lines.lcr %>% filter (bicycle >= bike_min)

#Create a new threshold for 20 trips min
bike_20 <- lines.inter %>% filter (bicycle >= 15)
nrow(bike_20)
plot(bike_20$geometry)
#create top all commutes
#trips.all.min <- 20
#lines.all.top <- lines.inter %>% filter (all >= trips.all.min)
```


```{r}
#check representativeness
nrow(bike.top) / nrow(lines.lcr)
sum(bike.top$all)/sum(lines.lcr$all)
```
3.7% of the bike top lines contain 29% of the total number of trips.

```{r}
#plot inter regional flows for bicycle
lwd = lines.inter$all / mean(lines.inter$all)/10
#create not driving flows
#create a percentage of cycling
bike.top$p_cycle <- bike.top$bicycle / bike.top$all *100
#create percent that does not drive
#lines.inter$percent.active <- 100 - (lines.inter$bicycle+lines.lcr$foot) / lines.inter$all *100
#create a percentage cycle
#lines.lcr$percent.cycle <- (lines.lcr$bicycle/ lines.lcr$all) *100
#plot the map
#Define line width for bike
lwd_bk = bike.top$all / mean(bike.top$all)
plot(bike.top["bicycle"], lwd = lwd)
```

  tm_borders(col = "grey", lwd = 0.3) + 
  tm_scale_bar(size = 0.3, width = 0.15 ,position = c("left", "bottom"))+
  tm_compass(type = "arrow", size = 0.7, position = c("right", "bottom"))+
  tm_layout(legend.position = c("right", "top"))

https://docs.ropensci.org/stplanr/articles/stplanr-od.html
https://github.com/Robinlovelace/geocompr/blob/master/code/12-desire.R

  tm_scale_bar(size = 0.3, width = 0.15,lwd= 0.3,
               position = c("left", "bottom")) +
  tm_compass(type = "arrow", size = 0.7, position = c("right", "bottom"))+

```{r}
lapply(bike.top, class)
```


```{r}
#calculate distance of top bikes 
bike.top$distance <- st_length(bike.top$bicycle)
summary(bike.top$bicycle)
```


```{r}
#Step 5
#plot with tmap top bike trajectories with min 3 trips
#tmap_mode("plot")
#bike.top = bike.top.all %>% 
#  arrange(bicycle) 
bike_all <- tm_shape(boundary) + tm_borders() + tm_fill("grey86")+
  tm_shape(bike.top) +
  tm_lines(palette = "plasma", breaks = c(3, 10, 20, 40, 50),
    lwd = "bicycle",
    #n=5,
    scale = 9,
    title.lwd = "Number of trips",
    alpha = 0.9,
    col = "bicycle",
    title = "Commutes by cycle",
    legend.lwd.show = FALSE) +
  tm_scale_bar(size=.5, width = 0.15,lwd= 0.3,
               position = c("left", "bottom"))+
   tm_compass(type = "arrow", size = 0.7, position = c("right", "bottom"))+
  tm_layout(
    legend.bg.alpha = 0.5,
    legend.bg.color = "white",
    legend.position = c("right", "top"),
    legend.text.size= 0.6,
    legend.stack="vertical")
bike_all
```

```{r}
tmap_save(bike_all, filename = "bike_all_flows_inv.jpeg", )
```


```{r}
summary(lines.inter$bicycle)
```

Using the same data, bike.top, create the street level network
```{r}
#Step 6 convert OD lines to street level lines
bikes_net <- line2route(bike.top, route_fun = route_cyclestreet)
bikes_net_quiet <- line2route(bike.top, plan = "quietest")
bike_balance <- line2route(bike.top, plan= "balanced")
plot(bikes_net_quiet$geometry)
```

the created network need to be joined to the bike.top. First, create an id column for bike.top
```{r}
#create id to join
bike.top <- bike.top %>% 
  mutate(id = row_number())
head(bike.top)
```

```{r}
#save as table drop geometry
#subset data keep table only no geo
bike_nogeo <- bike.top %>%
  st_drop_geometry() %>%
  select(id,geo_code1, geo_code2, all, bicycle)
bikes_net <- bikes_net %>%
  select(id)
#create for quiet routes
bikes_net_quiet <- bikes_net_quiet%>%
  select(id)
#create for balanced routes
bike_balance <- bike_balance %>%
  select(id)
```

```{r}
#join with the bikes_net
#dont know where this came from
#join data
#routes_commute = left_join(routes_commute, lines_commute)
#plot(routes_commute["bicycle"])
bike_network_all <- left_join(bikes_net, bike_nogeo)
plot(bike_network_all$geometry)
#try for quiet routes
bike_all_quiet <- left_join(bikes_net_quiet, bike_nogeo)
plot(bike_all_quiet$geometry)
#join for balanced routes
bike_balancenet <- left_join(bike_balance, bike_nogeo)
plot(bike_balancenet$geometry)
```


```{r}
#step 7
#create route network
#Route summarized with overline
#this results in a shape without any data, just lines
bikes_routesnet <- overline2(x = bike_network_all, "bicycle")
#create netowrk for quiet
bikes_routesquiet <- overline2(x= bike_all_quiet, "bicycle")
#create network for balanced
bikes_routesbalan <- overline2(x= bike_balancenet, "bicycle")
#plot
tm_shape(liv.shape)+tm_polygons()+
  tm_shape(bikes_routesbalan)+tm_lines(lwd = "bicycle", scale = 9, col = "bicycle" )
```

```{r}
summary(bikes_routesnet$bicycle)
summary(bikes_routesbalan$bicycle)
```


```{r}
#step 8
#create the same visualization as the OD map
#plot with tmap top bike trajectories with min 3 trips
bike_route <- tm_shape(boundary) + tm_borders() + tm_fill("grey86")+
  tm_shape(bikes_routesbalan) +
  tm_lines(palette = "plasma", breaks = c(3,15,30,50,100, 400),
    lwd = "bicycle",
    n=5,
    scale = 9,
    title.lwd = "Number of trips",
    alpha = 0.9,
    col = "bicycle",
    title = "Commutes by cycle",
    legend.lwd.show = FALSE) +
  tm_scale_bar(size=.5, width = 0.15,lwd= 0.3,
               position = c("left", "bottom"))+
   tm_compass(type = "arrow", size = 0.7, position = c("right", "bottom"))+
  tm_layout(
    legend.bg.alpha = 0.5,
    legend.bg.color = "white",
    legend.position = c("right", "top"),
    legend.text.size= 0.5,
    legend.stack="vertical")
bike_route
```

```{r}
tmap_save(bike_route, "route_cycle.jpeg")
```


```{r}
#Step 9
#project
bikes_rnet_pr = st_transform(bikes_routesnet, st_crs(boundary))
#project quiet routes
bikes_routes_quietpr = st_transform(bikes_routesquiet, st_crs(boundary))
#project balanced routes
bikes_balnetproj <- st_transform(bikes_routesbalan, st_crs(boundary))
#export 
st_write(bikes_balnetproj, "bike_top_balance.shp", driver =  "ESRI Shapefile")
```

###How to select where cycling commutes are promminent but that are not covered by cycling infrastructure

```{r}
#Step 10
#load cycling infrastructure
cycle <- st_read("cycle_all_proj.shp")
#calculate routes covered by cycle lanes
#diff <- st_sym_difference(cycle, bikes_balnetproj)
```

### create a regional map of flows
```{r}
summary(lines.inter$all)
nrow(lines.inter)
```

```{r}
#select top all flows the mean is 15.9 and max is 722
#define minimum trips for inter flows
top.all <- lines.inter %>% filter (all >= 100)
nrow(top.all)
```

```{r}
summary(top.all$all)
```


```{r}
#lwd = top.all$all / mean(top.$all)/10
#Create visualization
tm_shape(liv_shape)+tm_borders()+tm_fill()+
  tm_shape(top.all)+
  tm_lines(palette = "viridis", breaks = c(100,200,350,500,750),
    lwd = "all",
    scale = 8,
    title.lwd = "Number of trips",
    alpha = 0.9,
    col = "all",
    title = "Total commutes",
    legend.lwd.show = FALSE)+
    tm_scale_bar(size=.4, width = 0.15, position = c("left", "bottom")) +
  tm_layout(
    legend.bg.alpha = 0.5,
    legend.bg.color = "white",
    legend.position = c("right", "top"),
    legend.text.size= 0.5,
    legend.stack="horizontal")
```

```{r}
#Read boundary
boundary <- st_read("LCR_Boundary.shp")
```


```{r}
#Step 9
#re project
top.all.proj = st_transform(top.all, st_crs(boundary))
st_crs(top.all.proj)
 
st_write(top.all.proj, "top_all.shp", driver =  "ESRI Shapefile")
```


```{r}
#make one map with all flows smaller than 5
#select top bicycle flows the mean is 37 and max is 47
#define minimum trips
bike.min <- lines.inter %>% filter (bicycle <= 5  )
nrow(bike.min)
#nrow(lines.lcr)
```

```{r}
#Create map wth flows smaller than 5 trips. COmpletelet 
bike.min = bike.min %>% 
  arrange(bicycle) 
tm_shape(boundary) + tm_borders() +
  tm_shape(bike.top) +
  tm_lines(palette = "plasma", breaks = c(5, 10, 20, 40,50),
    lwd = "all",
    n=5,
    scale = 9,
    title.lwd = "Number of trips",
    alpha = 0.5,
    col = "bicycle",
    title = "Commutes by cycle",
    legend.lwd.show = TRUE) +
  tm_scale_bar(size=.4) +
  tm_layout(
    legend.bg.alpha = 0.5,
    legend.bg.color = "white",
    legend.position = c("right", "top"),
    legend.text.size= 0.6,
    legend.stack="horizontal")
```


tm_layout(
    legend.title.size = 1.5,
    legend.bg.alpha = 0.5,
    legend.bg.color = "white",
    legend.text.size = 0.8,
    legend.width=1.5,
    legend.stack="horizontal",
    legend.position = c("right", "top"))

```{r}
#Find commutes with trips less than 5
#select lowest bicycle flows the mean is 37 and max is 47
#define minimum trips for all flows
# <- lines.inter %>% filter (bicycle >= 2 & bicycle <= 4) %>%
#  count(bicycle)
#Define and count trips 1
min_trips <- lines.inter %>% filter (bicycle <= 2) %>%
  count(bicycle)
#Define trips bigger than 5
#min_trips <- lines.inter %>% filter (bicycle >= 5)
```

Total trips per trajectory for inter regional flows
2 trips = 782 flows
3 trips = 346 flows
4 trips = 178

```{r}
#all commutes minimum 5 trips
#plot with tmap
#tmap_mode("plot")
#min_trips = min_trips %>% 
 # arrange(bicycle)
tm_shape(boundary) + tm_borders() + tm_fill()+
  tm_shape(min_trips) +
  tm_lines(palette = "plasma", breaks = c(0,2,3,4,5),
    #lwd = "bicycle",
    #scale = 9,
    title.lwd = "Number of trips",
    alpha = 0.5,
    col = "bicycle",
    title = "Total comutes",
    legend.lwd.show = FALSE
  ) +
  tm_scale_bar() +
  tm_layout(
    legend.bg.alpha = 0.5,
    legend.bg.color = "white"
    )
```

```{r}
#plot of flows not made by car includes public transport, cycle, foot
#public transport is bus, rail, train, tram
#car includes passenger and driving?

```

#calculate distance for bike trips

```{r}
#calculate distance for cycle commutes
top300.car$Distance <-   st_length(top300.car)
top300.cycle$Distance <- st_length(top300.cycle)
```
