
```{r}
#install.packages("stplanr")
library(stplanr)
library(dplyr)
library(rgdal)
library(sf)
library(tidyverse)
library(pct)
library(tmap)
library(sp)
library(tmap)
library(ggplot2)
```

```{r}
rm (list=ls( ))
getwd()
```


```{r}
#Saving  my API token to R environment
mytoken = 'd6fee7f1acac5c7c'
Sys.setenv(CYCLESTREET = mytoken)
```

```{r}
#Load liv shape
boundary <- st_read("LCR_Boundary.shp")
```


```{r}
#How was the pct created. First, load centroids, area and OD pairs
#get centroids
centroids <- get_centroids_ew()
plot(centroids$geometry)
#load lcr shape
liv.shape <- pct_regions[pct_regions$region_name=="liverpool-city-region", ]
plot(liv.shape)
#plot a map
tm_shape(liv.shape)+tm_polygons()
#load centroids from region set crs
centroids = st_transform(centroids, st_crs(liv.shape))
#clip centroids of all EW using the LCR shape
centroids.lcr = centroids[liv.shape, ]
plot(centroids.lcr$geometry)
#get od data is the method to travel to work data
od = get_od(region = "liverpool-city-region")
#select od from LCR
od_liv = od[od$geo_code1 %in% centroids.lcr$msoa11cd, ]
head(od_liv)
#filter by destination in england and wales
#od_lcr_ew = od_liv[od_liv$geo_code2 %in% centroids$msoa11cd, ]
#filter by origin and destination in LCR
od_lcr = od_liv[od_liv$geo_code2 %in% centroids.lcr$msoa11cd, ]
#select top 50 od pairs
#od_lcr_top50 = od_lcr_ew %%
#  filter(all > 50)
#create od lines or desire lines- straight lines
lines_lcr = od2line(od_lcr, centroids)
plot(lines_lcr$geometry)
#street level lines with route quietest
street_lines <- line2route(lines_lcr[1:5, ], plan="quietest")
plot(street_lines$geometry)
```

```{r}
#calculate the distance of all trips in the LCR
#lines_lcr$Distance <- st_length(lines_lcr$geometry)
head(lines_lcr)
```


```{r}
?route_cyclestreet
```



#finish https://itsleeds.github.io/pct/articles/pct_training.html#test-code
#top 300 by car

#Top_n funtion help
#https://luisdva.github.io/rstats/Top-and-bottom-values-for-groups/

```{r}
#MSOA for inter regional commutes
#get data by MSOA
msoa <- get_pct_zones(region="liverpool-city-region", geography = "msoa")
#get lines by MSOA
lines_msoa <- get_pct_lines(region="liverpool-city-region", geography = "msoa")
plot(lines_msoa$geometry)
#select top 10 lines by bike
top10_bike_msoa <- lines_msoa%>% top_n(100, bicycle)
#plot maps
plot(msoa$geometry, axes=TRUE)
plot(top10_bike_msoa, add=TRUE, col = "red")
#select top 300 bike
top300.cycle <- lines_msoa%>% top_n(300,bicycle)
#select 300 top car
top300.car <- lines_msoa%>% top_n(300, car_driver)
#plot
plot(msoa$geometry, axes=TRUE)
plot(top300.cycle, add=TRUE, col = "red", lwd=top300.cycle$bicycle/ 
      mean(top300.cycle$bicycle))
```

```{r}
#plot cycling and car lines
plot(liv.shape$geometry, axes=TRUE)
plot(top300.cycle$geometry, add=TRUE, col="green")
plot(top300.car$geometry, add=TRUE, col="red")
```

We need to create inter regional flows too to run this lines
```{r}
#create desire lines for LCR
lines_lcr = od2line(od_lcr, centroids)
nrow(lines_lcr)
#create inter regional flows
lines.inter <- lines_lcr %>% filter(geo_code1 != geo_code2)
nrow(lines.inter)
#create intra regional
lines.intra <- lines_lcr %>% filter(geo_code1 == geo_code2)
```

```{r}
plot(lines.inter$geometry)
```


```{r}
#Calculate distance for all inter regional flows
lines.inter$Distance1 <- st_length(lines.inter$geometry)
#change to number 
lines.inter$Distance1 <- as.numeric(lines.inter$Distance1)
#Change to km
lines.inter$Distkm <- lines.inter$Distance1/1000
#lines.inter$Distance2 <- st_length(lines.inter)
head(lines.inter)
```

```{r}
summary(lines.inter$bicycle)
```


```{r}
#save shape for inter regional commutes
st_write(lines.inter, "lines_inter.shp", driver= "ESRI Shapefile")
```

```{r}
plot(lines_msoa$geometry)
```

```{r}
#Calculate distance for all inter regional flows
lines_msoa$Distance1 <- st_length(lines_msoa$geometry)
#change to number 
lines_msoa$Distance1 <- as.numeric(lines_msoa$Distance1)
#Change to km
lines_msoa$Distkm <- lines_msoa$Distance1/1000
#lines.inter$Distance2 <- st_length(lines.inter)
#subset file
lines_msoa <- lines_msoa[,c(1:3,10:13,88:90)]
head(lines_msoa)
```

```{r}
st_write(lines_msoa, "lines_msoa2.shp", driver="ESRI Shapefile")
```


```{r}
#select top bicycle flows the mean is 37 and max is 47
bike_min <- 3
#define minimum trips for inter flows
bike.top <- lines_msoa %>% filter (bicycle >= bike_min)
nrow(bike.top)
nrow(lines_msoa)
summary(bike.top$bicycle)
#define minimum trips for all flows
#bike.top.all  <- lines.lcr %>% filter (bicycle >= bike_min)
```

```{r}
tm_shape(boundary)+tm_borders()+tm_fill()+
  tm_shape(bike.top)+
  tm_lines(palette = "plasma", breaks = c(3,10,15, 20, 50),
    lwd = "bicycle",
    scale = 9,
    title.lwd = "Number of trips",
    alpha = 0.9,
    col = "bicycle",
    title = "Commutes by cycle",
    legend.lwd.show = FALSE)+
    tm_scale_bar(size=.4, width = 0.15, position = c("left", "bottom")) +
  tm_layout(
    legend.bg.alpha = 0.5,
    legend.bg.color = "white",
    legend.position = c("right", "top"),
    legend.text.size= 0.5,
    legend.stack="horizontal")
```



```{r}
#subset bike top
bike_top <- bike.top[,c(1:3,10:18)]
```

```{r}
#calculate distance of commute by bike
#top300.car$Distance <-   st_length(top300.car)
bike_top$Distance <- st_length(bike_top)
head(bike_top)
```

```{r}
#change data type units of cycle lane length to numeric (from units with m suffix)
bike_top$Distance <- as.numeric(bike_top$Distance)
bike_top$Distance <- bike_top$Distance/1000
head(bike_top)
```

```{r}
st_write(bike_top, "bike-distance.shp", driver="ESRI Shapefile")
```

```{r}
summary(bike_top$Distance)
```


#Plot trips and distance
```{r}
ggplot(data=bike_top)+
geom_line(aes(x=Distance, y=bicycle), color="cyan")+
  labs(x= "Distance (km)", y= "Bicycle trips")+
  theme_dark()
```


lines_msoa is the complete file
```{r}
#add percentage to cycle by MSOA
bike_top$pcycle <- (bike_top$bicycle / bike_top$all) * 100
plot(bike_top["pcycle"], lwd = bike_top$all/mean(bike_top$all), breaks=c(0,5,10,20,30,40,50))
```

Modifying PCT data to identify routes/roads of interest
Exercises from PCT workshop
```{r}
#download route network
network <- get_pct_rnet(region = "liverpool-city-region", geography = "msoa")
plot(liv.shape$geometry, axes= TRUE)
plot(network["bicycle"], add=TRUE)
#top 5street routes
top_street_cycle <- top_n(network, 4500, bicycle)
plot(liv.shape$geometry, axes=TRUE)
plot(top_street_cycle$geometry, col="red", add=TRUE)
#top 5 go dutch
top_street_godutch <- top_n(network,4500, dutch_slc)
plot(liv.shape$geometry, axes=TRUE)
plot(top_street_godutch$geometry, col="red", add=TRUE)
plot(top_street_cycle$geometry, col="green", add=TRUE)
```

```{r}
## Scenarios of change
#generate a godutch scenario
bike_top$euclidean_dist <- as.numeric(sf::st_length(isle_short))
isle_short$pcycle_dutch <- uptake_pct_godutch(
  distance=isle_short
)
```

```{r}
#Routing, this plots single line
route_bike <- route_osrm(bike_top)
```

```{r}
#plot
tm_shape(liv.shape) + tm_borders() +
tm_shape(route_bike)+ tm_lines(lwd = 5)
```

```{r}
#Load route network available at LSOA level
lcr_network <- get_pct_rnet(region = "liverpool-city-region", geography = "lsoa")
summary(lcr_network$bicycle)
plot(lcr_network$geometry)
```

```{r}
#clip lcr network with lcr region
liv_shape <- pct_regions[pct_regions$region_name=="liverpool-city-region", ]
st_crs(liv_shape)
lcr_net.clip <- lcr_network[liv_shape, ]
plot(lcr_net.clip$geometry)
```



```{r}
#top 10 route by bike
summary(lcr_net.clip$bicycle)
lcr.top10bike <- top_n(lcr_net.clip, 10, bicycle)
lct.top100bike <- top_n(lcr_network, 100, bicycle)
lcr.top500bike <- top_n(lcr_net.clip, 9000, bicycle)
top.bike <- top_n(lcr_net.clip, 4000, bicycle)
plot(liv.shape$geometry, axes=TRUE)
plot(lcr.top500bike$geometry, add=TRUE, col="red")
```

```{r}
#create the top 30 cycling routes in LCR
top.routes <- top_n(lcr_net.clip, 1000, bicycle)
summary(top.routes$bicycle) 
```


#calculate distance for both cycle and car
#calculate distance of commute by car
lcr.top500bike$distance <- st_length(lcr.top500bike$bicycle)


```{r}
#Route summarized with overline
#this results in a shape without any data, just lines
route_test<- overline(sl = network, attrib = "bicycle")
plot(route_test)
```

When saved as a shapefile it can be modified in qgis, better visualization.
```{r}
st_write(lcr.top500bike, "route_test.shp", driver = "ESRI Shapefile")
```


route

Use route as well
```{r}
#Map of the most used routes
route_cycle <- tm_shape(boundary)+tm_borders()+tm_fill()+
  tm_shape(lcr.top500bike)+
  tm_lines(palette = "plasma", breaks = c(0, 20, 50, 80,100,300),
    lwd = "bicycle",
    scale = 9,
    title.lwd = "Number of trips",
    alpha = 0.9,
    col = "bicycle",
    title = "Commutes by cycle",
    legend.lwd.show = FALSE)+
    tm_scale_bar(size=.4, width = 0.15, position = c("left", "bottom")) +
  tm_layout(
    legend.bg.alpha = 0.5,
    legend.bg.color = "white",
    legend.position = c("right", "top"),
    legend.text.size= 0.5,
    legend.stack="horizontal")
route_cycle
```

```{r}
tmap_save(route_cycle, filename = "route_cycle.jpeg")
```

```{r}
#try plotting both desire lines and cycling network in the same map
tm_shape(boundary) + tm_borders() + tm_fill("grey86")+
  tm_shape(bike.top) +
  tm_lines(col = "black") +
  
  tm_scale_bar(size=.5, width = 0.15,lwd= 0.3,
               position = c("left", "bottom"))+
   tm_compass(type = "arrow", size = 0.7, position = c("right", "bottom"))+
  tm_layout(
    legend.bg.alpha = 0.5,
    legend.bg.color = "white",
    legend.position = c("right", "top"),
    legend.text.size= 0.6,
    legend.stack="vertical")
```


```{r}
#to plot a network using routes fast
#load data for all routes
routes_commute <- get_pct_routes_fast(region = "liverpool-city-region", geography = "msoa")
lines_commute <- get_pct_lines(region = "liverpool-city-region", geography = "msoa")
```



```{r}
#subset data keep table only no geo
lines_commute <- lines_commute %>%
  st_drop_geometry() %>%
  select(id, bicycle, dutch_slc)
routes_commute <- routes_commute %>%
  select(id)
```

```{r}
#dont know where this came from
#join data
routes_commute = left_join(routes_commute, lines_commute)
plot(routes_commute["bicycle"])
```

```{r}
#create route network
rnet <- overline2(x=routes_commute, "bicycle")
tm_shape(liv.shape)+tm_polygons()+
  tm_shape(rnet)+tm_lines(lwd = "bicycle", scale = 9, col = "bicycle" )
```

#calculate distance for both cycle and car
#calculate distance of commute by car
isle_top_car$Distance <-   st_length(isle_top_car)

#to.place<- c(coords here)
#routes <- otp_plan(optcon = optcon,
fromPlace= lsoa(example),
tpPlace= to.place)

```{r}
boundary <- st_read("LCR_boundary.shp")
st_crs(boundary)
```

```{r}
#reproject to osgb36
rnet_proj = st_transform(rnet, st_crs(boundary))
lcr_network_proj = st_transform(lcr_network, st_crs(boundary))
st_crs(rnet_proj)
```


```{r}
#compare rnet and lcr_network
#save rnet
st_write(rnet_proj, "rnet_strplanr_proj.shp", driver = "ESRI Shapefile")
#save lcr_network
st_write(lcr_network_proj, "network_pct_proj.shp", driver="ESRI Shapefile")
```



Exercise at PCT workshop
```{r}
#Identify cycling routes street level and compare to the existent cycling infrastructure
lcr_network <- get_pct_rnet(region = "liverpool-city-region", geography = "msoa")
plot(lcr_network$geometry)
#load lcr zone
lcr_zone <- pct_regions[pct_regions$region_name=="liverpool-city-region", ]
lcr_poly <- get_pct_zones(region="liverpool-city-region", geography = "lsoa")
plot(lcr_zone$geometry, axes=TRUE)
plot(lcr_network$geometry, add=TRUE, col="red")

#top 10 route by bike
summary(lcr_network$bicycle)
lcr.top10bike <- top_n(lcr_network, 10, bicycle)
lct.top100bike <- top_n(lcr_network, 100, bicycle)
lcr.top500bike <- top_n(lcr_network, 500, bicycle)
plot(lcr_zone$geometry, axes=TRUE)
plot(lcr.top500bike$geometry, add=TRUE, col="red")


tm_shape(lcr.top500bike)+tm_lines(lcr.top500bike)
tm_shape(shape.wy)+tm_polygons()

setwd("C:/Users/PipeReds/DatAnywhere/SGVROJA2/Documents/PCT")
getwd()
#load road infrastructure
osm_cycle <- `osm-lines`
#set projection
osm_cycle = st_transform(osm_cycle, st_crs(lcr_zone))
#clip osm cycle to lcr zone
osm_lcr <- osm_cycle[lcr_zone, ]
#create top 20 routes

plot(lcr_zone$geometry, axes=TRUE)
plot(osm_lcr["roadtype"], add=TRUE, col="red")
#osm lcr subset to cycleway only or something
```

